manuscript_figures

### 2024-07-09 scripts/manuscript_figures.Rmd
updated habitat productivity index plots

# Now you can use this new 'Species' column in your ggplot
production_plot <- ggplot(PB_filtered, aes(x = YEAR, y = Production, fill = Species)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Year",  # X-axis label
       y = expression("Fish Production" ~ (g ~ m^-2 ~ yr^-1)),  # Y-axis label
       fill = "Species") +  # This should now correctly set the legend title
  scale_fill_viridis_d() +  # Use viridis colorblind-friendly palette
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12),  # Rotate x-axis labels if needed
        axis.title.x = element_text(vjust = -0.5, size = 14),  # Adjust x-axis title position
        axis.title.y = element_text(vjust = 1.5, size = 14),
        axis.text.y = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14)) +
  scale_y_continuous(limits = c(0, 4), expand = c(0, 0));production_plot  # Additional attempt to change legend title  # Use viridis colorblind-friendly palette

@@

# Log transform HPI 
HPI_calculated <- HPI_calculated %>%
  mutate(log_HPI = log(HPI + 1.1))

# Define the color palette
my_palette <- brewer.pal(n = 6, name = "Paired")  # Adjust 'n' as per the number of years in your data

# Plotting code
hpi.plot <- ggplot(HPI_calculated, aes(x = SITE, y = log_HPI, fill = factor(YEAR))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Site", 
       y = "Log Habitat Productivity Index", 
       fill = "Year") +
  scale_fill_manual(values = my_palette) +  # Use the colorblind-friendly palette
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
        axis.text.y = element_text(size = 18),
        axis.title.y = element_text(size = 18),
        axis.title.x = element_text(size = 18),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18)) +
  scale_y_continuous(expand = expansion(mult = c(0, .05)),
                     labels = scales::label_number(accuracy = 0.1)) +
  coord_cartesian(clip = "off");hpi.plot

# Plot two plots together using patchwork for production and HPI
```{r}
# Combine the plots with patchwork for all years of kelp and rate of change
production.hpi <- production_plot | hpi.plot +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 20, face = "bold"));production.hpi

# Save the combined plot
ggsave("output/figures/production.HPI.png", production.hpi, width = 20, height = 10, units = "in", dpi = 300)

>>>>

# Now you can use this new 'Species' column in your ggplot
production_plot <- ggplot(PB_filtered, aes(x = YEAR, y = Production, fill = Species)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Year",  # X-axis label
       y = expression("Fish Production" ~ (g ~ m^-2 ~ yr^-1)),  # Y-axis label
       fill = "Species") +  # This should now correctly set the legend title
  scale_fill_viridis_d() +  # Use viridis colorblind-friendly palette
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),  # Rotate x-axis labels if needed
        axis.title.x = element_text(vjust = -0.5, size = 18),  # Adjust x-axis title position
        axis.title.y = element_text(vjust = 1.5, size = 18),
        axis.text.y = element_text(size = 18),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18)) +
  scale_y_continuous(limits = c(0, 4), expand = c(0, 0));production_plot  # Additional attempt to change legend title  # Use viridis colorblind-friendly palette

@@

# Add log_HPI column to the dataframe
HPI_calculated <- HPI_calculated %>%
  mutate(log_HPI = log(HPI))

# Summarize by year and site for plot
HPI_summary <- HPI_calculated %>%
  group_by(YEAR, SITE) %>%
  summarise(
    mean_HPI = mean(HPI, na.rm = TRUE),
    se_HPI = sd(HPI, na.rm = TRUE) / sqrt(n()),
    mean_log_HPI = mean(log_HPI, na.rm = TRUE),
    se_log_HPI = sd(log_HPI, na.rm = TRUE) / sqrt(n()),
    .groups = 'drop'
  )

# Create the plot with nonlogged values
hpi.plot.nonlog <- ggplot(HPI_summary, aes(x = factor(YEAR), y = mean_HPI, group = SITE)) +
  geom_line() +
  geom_point(aes(shape = SITE), size = 3) +
  geom_errorbar(aes(ymin = mean_HPI - se_HPI, ymax = mean_HPI + se_HPI), width = 0.2) +
  labs(x = "Year", 
       y = "Habitat Productivity Index", 
       shape = "Site") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
        axis.text.y = element_text(size = 18),
        axis.title = element_text(size = 18),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18)) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.05)),
                     labels = scales::label_number(accuracy = 0.1));hpi.plot.nonlog

# Create the plot with log_HPI values
hpi_plot_log <- ggplot(HPI_summary, aes(x = factor(YEAR), y = mean_log_HPI, group = SITE)) +
  geom_line() +
  geom_point(aes(shape = SITE), size = 3) +
  geom_errorbar(aes(ymin = mean_log_HPI - se_log_HPI, ymax = mean_log_HPI + se_log_HPI), width = 0.2) +
  labs(x = "Year", 
       y = "Log Habitat Productivity Index", 
       shape = "Site") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
        axis.text.y = element_text(size = 18),
        axis.title = element_text(size = 18),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18)) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.05)),
                     labels = scales::label_number(accuracy = 0.1));hpi_plot_log

# Define the color palette
#my_palette <- brewer.pal(n = 6, name = "Paired")  # Adjust 'n' as per the number of years in your data

# Test for normality and perform appropriate statistical test
# Check for normality using Shapiro-Wilk test for each site
shapiro_test_fort_wetherill <- shapiro.test(HPI_calculated$HPI[HPI_calculated$SITE == "Fort Wetherill"])
shapiro_test_kings_beach <- shapiro.test(HPI_calculated$HPI[HPI_calculated$SITE == "King's Beach"])

# Print results of normality tests
print(shapiro_test_fort_wetherill)
print(shapiro_test_kings_beach)

# Randomize the YEAR column
set.seed(123)  # Set seed for reproducibility
HPI_calculated$YEAR <- sample(HPI_calculated$YEAR)

# Perform the Mann-Whitney U test on logged HPI values
wilcox_test_log_result <- wilcox.test(HPI ~ SITE, data = HPI_calculated)

# Print the result of the Wilcoxon rank-sum test
print(wilcox_test_log_result)
```
Wilcoxon rank sum exact test

data:  HPI by SITE
W = 527, p-value = 0.03417
alternative hypothesis: true location shift is not equal to 0

# Plot two plots together using patchwork for production and HPI
```{r}
# Combine the plots with patchwork for all years of kelp and rate of change
production.hpi.nonlog <- production_plot | hpi.plot.nonlog +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 20, face = "bold"));production.hpi.nonlog

# Save the combined plot
ggsave("output/figures/production.HPI.nonlog.png", production.hpi.nonlog, width = 20, height = 10, units = "in", dpi = 300)

# Combine the plots with patchwork for all years of kelp and rate of change
production.hpi.log <- production_plot | hpi_plot_log +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 20, face = "bold"));production.hpi.log

# Save the combined plot
ggsave("output/figures/production.HPI.log.png", production.hpi.log, width = 20, height = 10, units = "in", dpi = 300)
